---
title: "Untitled"
author: "Antoine Beauchamp"
date: '2023-06-28'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Initialization

```{r packages}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(RColorBrewer))
suppressPackageStartupMessages(library(viridis))
suppressPackageStartupMessages(library(viridisLite))
suppressPackageStartupMessages(library(SNFtool))
suppressPackageStartupMessages(library(mvtnorm))
suppressPackageStartupMessages(library(umap))
```

```{r environment}
SRCPATH <- Sys.getenv("SRCPATH")
PROJECTPATH <- Sys.getenv("PROJECTPATH")
```

```{r functions}
# compute_cluster_delta <- function(x, clusters, normalize = TRUE) {
#   
#   x[lower.tri(x, diag = TRUE)] <- NA
#   
#   x <- x %>% 
#     as_tibble(rownames = "ID1") %>% 
#     pivot_longer(cols = -ID1, 
#                  names_to = "ID2", 
#                  values_to = "similarity") %>% 
#     filter(!is.na(similarity)) %>% 
#     left_join(clusters, by = c("ID1" = "ID")) %>% 
#     rename(k1 = k) %>% 
#     left_join(clusters, by = c("ID2" = "ID")) %>% 
#     rename(k2 = k) %>% 
#     mutate(within = ifelse(k1 == k2, TRUE, FALSE))
#   
#   mu_within <- x %>% 
#     filter(within) %>% 
#     pull(similarity) %>% 
#     mean()
#   
#   mu_without <- x %>% 
#     filter(!within) %>% 
#     pull(similarity) %>% 
#     mean()
#   
#   if (normalize) {
#     sigma <- sd(x[["similarity"]])
#   } else {
#     sigma <- 1
#   }
#   
#   delta <- (mu_within - mu_without)/sigma
#   
#   return(delta)
#   
# }

# permute_cluster_delta <- function(x, clusters, npermutations = 1000, start = 1) {
#   delta <- numeric(npermutations)
#   for (p in start:npermutations) {
#     set.seed(p)
#     clusters[["k"]] <- sample(x = clusters[["k"]], size = nrow(clusters), replace = FALSE)
#     delta[p] <- compute_cluster_delta(x = x, clusters = clusters)
#   }
#   return(delta)
# }

#' Run similarity network fusion (SNF)
#'
#' @param x1 (matrix) Input matrix
#' @param x2 (matrix) Input matrix
#' @param metric (character scalar) Distance metric used to compute the 
#' affinity matrices.
#' @param K (numeric scalar) Number of nearest-neighbours used to 
#' compute the SNF affinity matrices. 
#' @param sigma (numeric scalar) Variance for the local model in the 
#' SNF affinity matrices.
#' @param t (numeric scalar) Number of iterations for the diffusion
#' process in SNF.
#' @param outfile (character scalar) Path to file in which to save
#' affinity matrix.
#'
#' @return (matrix) SNF affinity matrix.
# similarity_network <- function(x1, x2, metric = "correlation", K = 10,
#                                sigma = 0.5, t = 20, outfile = NULL){
#   
#   if (metric == "correlation") { 
#     d1 <- (1-cor(t(x1)))
#     d2 <- (1-cor(t(x2)))
#   } else if (metric == "euclidean") {
#     d1 <- (dist2(as.matrix(x1), as.matrix(x1)))^(1/2)
#     d2 <- (dist2(as.matrix(x2), as.matrix(x2)))^(1/2)
#   } else {
#     stop(paste("Argument metric must be one of {correlation, euclidean}:", metric))
#   }
#   
#   W1 <- affinityMatrix(d1, K = K, sigma = sigma)
#   W2 <- affinityMatrix(d2, K = K, sigma = sigma)
#   
#   W <- SNF(list(W1, W2), K = K, t = t)
#   
#   if (!is.null(outfile)){
#     data.table::fwrite(x = as_tibble(W), file = outfile)
#   }
#   
#   return(W)
#   
# }

#' Create clusters from SNF affinity matrix
#'
#' @param W (matrix) SNF affinity matrix.
#' @param nk (numeric scalar) Maximum number of clusters to use in 
#' clustering.
#' @param outfile (character scalar) Optional path to .csv file in 
#' which to save cluster assignments.
#'
#' @return (data.frame) Cluster assignments.
create_clusters <- function(W, nk = 10, outfile = NULL) {
  
  for(k in 2:nk) {
    group <- spectralClustering(affinity = W, K = k)
    group_name <- paste0("nk", k)
    assign(group_name, group)
    if (k == 2) {
      if (is.null(rownames(W))) {
        ids <- as.character(1:nrow(W))
      } else {
        ids <- rownames(W)
      }
      all_clusters <- data.frame(ids, group, stringsAsFactors = F)
      colnames(all_clusters) <- c("ID", group_name)
    } else {
      group <- data.frame(group)
      colnames(group) <- group_name
      all_clusters <- cbind(all_clusters, group)
    }
  }
  
  if (!is.null(outfile)) {
    write.csv(x = all_clusters, file = outfile, row.names = FALSE)
  }
  
  return(all_clusters)
  
}

estimate_cluster_metrics <- function (W, NUMC = 2:5){
  if (min(NUMC) == 1) {
    warning("Note that we always assume there are more than one cluster.")
    NUMC <- NUMC[NUMC > 1]
  }
  W <- (W + t(W))/2
  diag(W) <- 0
  if (length(NUMC) <= 0) {
    warning(paste("Invalid NUMC provided, must be an integer vector",
                  "with atleast one other number than 1.", "Using default NUMC=c(2,3,4,5)",
                  sep = ""))
    NUMC <- 2:5
  }
  degs <- rowSums(W)
  degs[degs == 0] <- .Machine$double.eps
  D <- diag(degs)
  L <- D - W
  Di <- diag(1/sqrt(degs))
  L <- Di %*% L %*% Di
  eigs <- eigen(L)
  eigs_order <- sort(eigs$values, index.return = T)$ix
  eigs$values <- eigs$values[eigs_order]
  eigs$vectors <- eigs$vectors[, eigs_order]
  eigengap <- abs(diff(eigs$values))
  quality <- list()
  for (c_index in 1:length(NUMC)) {
    ck <- NUMC[c_index]
    UU <- eigs$vectors[, 1:ck]
    EigenvectorsDiscrete <- SNFtool:::.discretisation(UU)[[1]]
    EigenVectors <- EigenvectorsDiscrete^2
    temp1 <- EigenVectors[do.call(order, lapply(1:ncol(EigenVectors),
                                                function(i) EigenVectors[, i])), ]
    temp1 <- t(apply(temp1, 1, sort, TRUE))
    quality[[c_index]] <- (1 - eigs$values[ck + 1])/(1 -
                                                       eigs$values[ck]) * sum(sum(diag(1/(temp1[, 1] + .Machine$double.eps)) %*%
                                                                                    temp1[, 1:max(2, ck - 1)]))
  }
  
  out <- tibble(nk = NUMC,
                eigengap = eigengap[NUMC],
                rotation = unlist(quality))
  
  return(out)
}

similarity_heatmap <- function(x, clusters, threshold = NULL) {
  
  x <- x %>% 
    as_tibble(rownames = "ID1") %>% 
    pivot_longer(cols = -ID1, names_to = "ID2", values_to = "similarity") %>% 
    left_join(clusters, by = c("ID1" = "ID")) %>% 
    rename(k1 = k) %>% 
    left_join(clusters, by = c("ID2" = "ID")) %>% 
    rename(k2 = k)
  
  lvls <- x %>% 
    select(ID1, k1) %>% 
    distinct() %>% 
    arrange(k1) %>% 
    pull(ID1)
  
  x <- x %>% 
    mutate(ID1 = factor(ID1, levels = lvls),
           ID2 = factor(ID2, levels = lvls),
           similarity = ifelse(ID1 == ID2, NA, similarity))
  
  if (!is.null(threshold)) {
    x <- x %>% 
      mutate(similarity = ifelse(similarity >= threshold, threshold, similarity))
  }
  
  p <- ggplot(x, aes(x = ID1, y = ID2, fill = similarity)) +
    geom_tile() + 
    scale_fill_viridis() + 
    theme(axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank())
  
  return(p)
  
}

umap_scatter <- function(u, clusters) {
  
  df_u <- u[["layout"]]
  colnames(df_u) <- c("x1", "x2")
  df_u <- as_tibble(df_u)
  df_u[["ID"]] <- rownames(u[["data"]])
  
  df_u <- df_u %>% 
    inner_join(clusters, by = "ID") %>% 
    mutate(k = factor(k))
  
  p <- ggplot(df_u, 
              aes(x = x1, y = x2, col = k)) + 
    geom_point()
  
  return(p)
  
}

plot_eigengap <- function(x, nk_max = 10) {
  p <- x %>% 
    estimate_cluster_metrics(NUMC = 2:nk_max) %>% 
    ggplot(aes(x = nk, y = eigengap)) + 
    geom_hline(yintercept = 0,
               linetype = "dashed",
               col = "grey50") + 
    geom_line() + 
    geom_point() +
    scale_x_continuous(breaks = seq(0, nk_max, by = 4)) + 
    labs(y = "Eigengap") +
    theme_bw()
  
  return(p)
}
```

```{r}
output_dir <- file.path(PROJECTPATH, "analyses", "auxiliary", "outputs", "cluster_simulations")
if (!file.exists(output_dir)) {
  dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)
}
```


```{r import}
human_pipeline_dir <- file.path(PROJECTPATH, "data/human/derivatives/v2/700/")
human_clusters_dir <- file.path(human_pipeline_dir, "clusters", "resolution_3.0")
human_es_dir <- file.path(human_pipeline_dir, "effect_sizes", "resolution_3.0")

human_clusters <- "clusters.csv"
human_clusters <- file.path(human_clusters_dir, human_clusters)
df_human_clusters <- read_csv(human_clusters, show_col_types = FALSE)

human_affinity <- "affinity.csv"
human_affinity <- file.path(human_clusters_dir, human_affinity)
df_human_affinity <- read_csv(human_affinity, show_col_types = FALSE)

human_es <- "effect_sizes.csv"

human_es_rel <- file.path(human_es_dir, "relative", human_es)
df_human_es_rel <- as_tibble(data.table::fread(human_es_rel, header = TRUE))

# human_es_abs <- file.path(human_es_dir, "absolute", human_es)
# df_human_es_abs <- as_tibble(data.table::fread(human_es_abs, header = TRUE))
```

# Human imaging data

```{r human-affinity}
df_human_clusters_nk <- df_human_clusters %>% 
  select(ID, k = nk2)

mat_human_affinity <- as.matrix(df_human_affinity)
rownames(mat_human_affinity) <- colnames(mat_human_affinity)

similarity_heatmap(x = mat_human_affinity, 
                   clusters = df_human_clusters_nk,
                   threshold = 0.015)
```


```{r human-cluster-delta}
compute_cluster_delta(x = mat_human_affinity, 
                      clusters = df_human_clusters_nk)
```

```{r human-eigengap}
plot_eigengap(x = mat_human_affinity, nk_max = 10)
```

```{r human-es-mat}
mat_es_rel <- df_human_es_rel %>% 
  select(-file) %>% 
  as.matrix()

mat_es_abs <- df_human_es_abs %>% 
  select(-file) %>% 
  as.matrix()

rownames(mat_es_rel) <- df_human_es_rel[["file"]]
rownames(mat_es_abs) <- df_human_es_abs[["file"]]
```

```{r human-umap}
umap_seed <- 2
umap_components <- 2

es_rel_umap <- umap(mat_es_rel, 
                    n_components = umap_components, 
                    random_state = umap_seed)

umap_scatter(u = es_rel_umap, 
             clusters = df_human_clusters_nk)
```


# Gaussian noise simulations

## Two-dimensional examples

### Uncorrelated Gaussian distribution

```{r gaussian-uncorrelated}
nvars <- 2
nobs <- 10000

mat <- rmvnorm(n = nobs, mean = rep(0, nvars), sigma = diag(nvars))
colnames(mat) <- c("x", "y")

ggplot(as_tibble(mat),
       aes(x = x, y = y)) + 
  geom_hex() + 
  coord_fixed() + 
  scale_fill_viridis() + 
  theme_bw()
```

```{r}
df <- tibble(x = rnorm(n = 1000), 
             y = rnorm(n = 1000, mean = 2))

df %>% 
  pivot_longer(cols = everything(), names_to = "dist", values_to = "x") %>% 
  ggplot(aes(x = x, fill = dist, col = dist)) + 
  geom_density(alpha = 0.5)
```



### Mixture of uncorrelated Gaussians

```{r gaussian-mixture}
nvars <- 10
nobs <- 1000

mat_1 <- rmvnorm(n = nobs/2, mean = rep(0, nvars), sigma = diag(nvars))
mat_2 <- rmvnorm(n = nobs/2, mean = c(1, rep(0, nvars-1)), sigma = diag(nvars))
mat <- rbind(mat_1, mat_2)
colnames(mat) <- paste0("x", 1:nvars)

# ggplot(as_tibble(mat), 
# aes(x = x, y = y)) + 
# geom_hex() + 
# coord_fixed() + 
# scale_fill_viridis() + 
# theme_bw()
```

```{r}
umap_out <- umap(mat, 
                 n_components = 2, 
                 random_state = 2)

df <- umap_out[["layout"]]
colnames(df) <- c("x", "y")
df <- as_tibble(df)
df$label <- rep(c(1, 2), each = nobs/2)

ggplot(df, aes(x = x, y = y, col = factor(label))) + 
  geom_point(alpha = 0.8)
```



### Correlated Gaussian distribution

```{r gaussian-correlated}
nvars <- 2
nobs <- 10000

sigma_ii <- 1.0
sigma_ij <- 0.9
sigma <- rbind(c(sigma_ii, sigma_ij), 
               c(sigma_ij, sigma_ii))
mat <- rmvnorm(n = nobs, mean = rep(0, nvars), sigma = sigma)
colnames(mat) <- c("x", "y")

ggplot(as_tibble(mat),
       aes(x = x, y = y)) + 
  geom_hex() + 
  coord_fixed() + 
  scale_fill_viridis() + 
  theme_bw()
```

```{r gaussian-cor-mat}
cor(mat)
```

```{r gaussian-cor-lm}
summary(lm(y ~ x, data = as_tibble(mat)))
```

So given one Gaussian variable, we can generate a highly correlated second variable using a linear model. The residual standard error needs to be around 0.5. 

```{r gaussian-cor-sim}
x <- rnorm(n = nobs)
y <- x + rnorm(n = nobs, sd = 0.5)

tibble(x = x, y = y) %>% 
  ggplot(aes(x = x, y = y)) + 
  geom_hex() +
  coord_fixed() + 
  scale_fill_viridis() + 
  theme_bw()
```

This will allow us to generate a correlated pair of images. 

## Simulations at image dimension

### Uncorrelated image pairs

```{r}
nobs <- nrow(mat_es_rel)
nvars <- ncol(mat_es_rel)

mat_sim_rel_nocor <- matrix(data = rnorm(n = nobs*nvars), nrow = nobs, ncol = nvars)
mat_sim_abs_nocor <- matrix(data = rnorm(n = nobs*nvars), nrow = nobs, ncol = nvars)

rownames(mat_sim_rel_nocor) <- rownames(mat_es_rel)
rownames(mat_sim_abs_nocor) <- rownames(mat_es_abs)

affinity_sim_nocor <- similarity_network(x1 = mat_sim_rel_nocor, x2 = mat_sim_abs_nocor)
clusters_sim_nocor <- create_clusters(W = affinity_sim_nocor)
```

```{r}
clusters_sim_nocor_nk <- select(clusters_sim_nocor, ID, k = nk2)
similarity_heatmap(x = affinity_sim_nocor, 
                   clusters = clusters_sim_nocor_nk,
                   threshold = 0.015)
```

```{r}
plot_eigengap(x = affinity_sim_nocor, nk_max = 10)
```

```{r}
umap_seed <- 2
umap_components = 2

sim_nocor_umap <- umap(mat_sim_rel_nocor, 
                       n_components = umap_components, 
                       random_state = umap_seed)

umap_scatter(u = sim_nocor_umap, 
             clusters = clusters_sim_nocor_nk)
```

```{r}
rm(list = ls()[str_detect(ls(), "_nocor")])
gc()
```



### Correlated image pairs

```{r}
es_cor <- numeric(nrow(mat_es_rel))
for (i in 1:length(es_cor)) {
  es_cor[i] <- cor(mat_es_rel[i,], mat_es_abs[i,])
}

tibble(x = es_cor) %>% 
  ggplot(aes(x = x)) + 
  geom_histogram()
```

Absolute and relative images correlated above 0.9

```{r}
nobs <- nrow(mat_es_rel)
nvars <- ncol(mat_es_rel)

mat_sim_rel_cor <- matrix(data = rnorm(n = nobs*nvars), nrow = nobs, ncol = nvars)
mat_sim_abs_cor <- matrix(data = 0, nrow = nobs, ncol = nvars)
for (i in 1:nobs) {
  mat_sim_abs_cor[i,] <- mat_sim_rel_cor[i,] + rnorm(n = nvars, sd = 0.5)
}

rownames(mat_sim_rel_cor) <- rownames(mat_es_rel)
rownames(mat_sim_abs_cor) <- rownames(mat_es_abs)

affinity_sim_cor <- similarity_network(x1 = mat_sim_rel_cor, x2 = mat_sim_abs_cor)
clusters_sim_cor <- create_clusters(W = affinity_sim_cor)
```

```{r}
clusters_sim_cor_nk <- select(clusters_sim_cor, ID, k = nk2)
similarity_heatmap(x = affinity_sim_cor, 
                   clusters = clusters_sim_cor_nk,
                   threshold = 0.015)
```

```{r}
compute_cluster_delta(x = affinity_sim_cor, clusters = clusters_sim_cor_nk, normalize = TRUE)
```

```{r}
plot_eigengap(x = affinity_sim_cor, nk_max = 10)
```

```{r}
umap_seed <- 2
umap_components = 2

sim_cor_umap <- umap(mat_sim_rel_cor, 
                     n_components = umap_components, 
                     random_state = umap_seed)

umap_scatter(u = sim_cor_umap, 
             clusters = clusters_sim_cor_nk)
```

```{r}
rm(list = ls()[str_detect(ls(), "_cor")])
gc()
```


### Single feature affinity matrix


```{r}
df_human_clusters_nk <- df_human_clusters %>% 
  select(ID, k = nk2)

# Number of observations per cluster
nobs_nk2 <- df_human_clusters_nk %>% 
  group_by(k) %>% 
  count() %>% 
  pull(n)

# Total number of observations
nobs <- sum(nobs_nk2)
# nvars <- ncol(df_human_es_rel)

# Number of variables
nvars <- ncol(df_human_es_rel)

niter <- 10
df_grid <- expand_grid(delta = seq(0, 10, by = 0.5),
                       seed = 1:niter)
results <- vector(mode = "list", length = nrow(df_grid))
# for (i in 1:length(results)) {

  i <- 1
    
  # print(paste("i:", i, "of", length(results)))
  # 
  # seed <- df_grid[[i, "seed"]]
  # delta <- df_grid[[i, "delta"]]
  
  seed <- 1
  delta <- 100
  
  # Generate normal distributions
  set.seed(seed)

  x1 <- matrix(rnorm(nobs_nk2[1]*nvars), nrow = nobs_nk2[1], ncol = nvars)
  x2_1 <- matrix(rnorm(nobs_nk2[2]*(nvars-1)), nrow = nobs_nk2[2], ncol = nvars-1)
  x2_2 <- rnorm(nobs_nk2[2], mean = delta)
  x2 <- cbind(x2_1, x2_2)
  x <- rbind(x1, x2)
  rownames(x) <- paste0("p", 1:nobs)
  
  # x1 <- rmvnorm(n = nobs_nk2[1], mean = rep(0, nvars), sigma = diag(nvars))
  # x2 <- rmvnorm(n = nobs_nk2[2], mean = c(delta, rep(0, nvars-1)), sigma = diag(nvars))
  # x <- rbind(x1, x2)

  # UMAP projection  
  umap_out <- umap(x, n_components = 2, 
                   random_state = seed)
  
  # Distance matrix
  d <- 1 - cor(t(x))
  
  # Free up memory
  rm(list = c("x1", "x2", "x")); gc()
  
  # Affinity matrix
  W <- affinityMatrix(d, K = 10, sigma = 0.5)
  
  # Run SNF because it seems to apply some normalization?
  W <- SNF(list(W, W), K = 10, t = 20)
  
  # Estimate clustering metrics
  metrics <- estimate_cluster_metrics(W = W, NUMC = 2:10)
  
  results[[i]] <- metrics %>% 
    mutate(seed = seed, delta = delta)
  
  clusters <- create_clusters(W = W, nk = 2)
  
  umap_x <- umap_out$layout
  colnames(umap_x) <- paste0("x", 1:2)
  p_umap <- umap_x %>% 
    as_tibble() %>% 
    bind_cols(clusters) %>% 
    ggplot(aes(x = x1, y = x2, col = factor(nk2))) + 
    geom_point() + 
    coord_equal() +
    labs(color = "Cluster",
         title = paste0("Delta: ", delta, "; Seed: ", seed)) + 
    theme_bw()
  
  outfile_umap <- paste("umap_delta", delta, "seed", seed, sep = "_")
  outfile_umap <- paste0(outfile_umap, ".pdf")
  outfile_umap <- file.path(output_dir, outfile_umap)
  pdf(file = outfile_umap,
      width = unit(8, "inch"),
      height = unit(8, "inch"))
  print(p_umap)
  dev.off()
  
# }
```



```{r}
df_metrics <- bind_rows(results)
ggplot(df_metrics, aes(x = nk, y = eigengap, col = factor(seed), group = factor(seed))) + 
  geom_line() + 
  geom_point() + 
  facet_wrap(~delta)
```



```{r}
# Create clusters from affinity matrix
clusters <- create_clusters(W = W, nk = 10)

clusters_nk <- clusters %>% 
  select(ID, k = nk2)

x_umap <- umap(x, n_components = 2, random_state = 2)

umap_scatter(u = x_umap, 
             clusters = clusters_nk)
```

```{r}
estimate_cluster_metrics(W = test, NUMC = 2:10)
```


```{r}
plot_eigengap(x = test, nk_max = 10)
```



```{r}
umap_seed <- 2
umap_components <- 2

mat_umap <- umap(mat, 
                 n_components = 2, 
                 random_state = 2)

umap_scatter(u = mat_umap, 
             clusters = clusters_nk)
```


```{r}
create_clusters <- function(W, nk = 10, outfile = NULL) {
  
  for(k in 2:nk) {
    group <- spectralClustering(affinity = W, K = k)
    group_name <- paste0("nk", k)
    assign(group_name, group)
    if (k == 2) {
      if (is.null(rownames(W))) {
        ids <- as.character(1:nrow(W))
      } else {
        ids <- rownames(W)
      }
      all_clusters <- data.frame(ids, group, stringsAsFactors = F)
      colnames(all_clusters) <- c("ID", group_name)
    } else {
      group <- data.frame(group)
      colnames(group) <- group_name
      all_clusters <- cbind(all_clusters, group)
    }
  }
  
  if (!is.null(outfile)) {
    write.csv(x = all_clusters, file = outfile, row.names = FALSE)
  }
  
  return(all_clusters)
  
}
```



```{r}
similarity_network <- function(x1, x2, metric = "correlation", K = 10,
                               sigma = 0.5, t = 20, outfile = NULL){
  
  if (metric == "correlation") {
    d1 <- (1-cor(t(x1)))
    d2 <- (1-cor(t(x2)))
  } else if (metric == "euclidean") {
    d1 <- (dist2(as.matrix(x1), as.matrix(x1)))^(1/2)
    d2 <- (dist2(as.matrix(x2), as.matrix(x2)))^(1/2)
  } else {
    stop(paste("Argument metric must be one of {correlation, euclidean}:", metric))
  }
  
  W1 <- affinityMatrix(d1, K = K, sigma = sigma)
  W2 <- affinityMatrix(d2, K = K, sigma = sigma)
  
  W <- SNF(list(W1, W2), K = K, t = t)
  
  if (!is.null(outfile)){
    data.table::fwrite(x = as_tibble(W), file = outfile)
  }
  
  return(W)
  
}
```



```{r}
mat_sim <- matrix(data = rnorm(n = nobs*nvars), nrow = nobs, ncol = nvars)
mat_sim_abs_nocor <- matrix(data = rnorm(n = nobs*nvars), nrow = nobs, ncol = nvars)

rownames(mat_sim_rel_nocor) <- rownames(mat_es_rel)
rownames(mat_sim_abs_nocor) <- rownames(mat_es_abs)

affinity_sim_nocor <- similarity_network(x1 = mat_sim_rel_nocor, x2 = mat_sim_abs_nocor)
clusters_sim_nocor <- create_clusters(W = affinity_sim_nocor)
```


## Lower dimensional

### With signal

```{r}
mu_1 <- rep(0, nvars)
mu_2 <- c(5, rep(0, nvars-1))

mat_gaussian_1_1 <- rmvnorm(n = nobs/2, mean = mu_1, sigma = diag(nvars))
mat_gaussian_1_2 <- rmvnorm(n = nobs/2, mean = mu_2, sigma = diag(nvars))
mat_gaussian_1 <- rbind(mat_gaussian_1_1, mat_gaussian_1_2)

mat_gaussian_2_1 <- rmvnorm(n = nobs/2, mean = mu_1, sigma = diag(nvars))
mat_gaussian_2_2 <- rmvnorm(n = nobs/2, mean = mu_2, sigma = diag(nvars))
mat_gaussian_2 <- rbind(mat_gaussian_2_1, mat_gaussian_2_2)

rownames(mat_gaussian_1) <- human_es[["file"]]
rownames(mat_gaussian_2) <- human_es[["file"]]

gaussian_affinity <- similarity_network(x1 = mat_gaussian_1, x2 = mat_gaussian_2,
                                        metric = "euclidean", K = 10, sigma = 0.5, t = 20)

gaussian_clusters <- create_clusters(W = gaussian_affinity, nk = 10)
```

```{r}
sqrt(sum((mu_1 - mu_2)^2))
```
```{r}
df_clusters_true <- tibble(ID = human_es[["file"]],
                           k_true = rep(1:2, each = nobs/2))

gaussian_clusters_nk <- gaussian_clusters %>% 
  select(ID, k = nk2)

df_clusters_compare <- inner_join(df_clusters_true, 
                                  gaussian_clusters_nk, 
                                  by = "ID")

noverlap <- df_clusters_compare %>% 
  mutate(overlap = ifelse(k_true == k, TRUE, FALSE)) %>% 
  pull(overlap) %>% 
  sum()

noverlap/nrow(df_clusters_compare)  
```


```{r}
df_gaussian_affinity_nk <- gaussian_affinity %>% 
  as_tibble(rownames = "ID1") %>% 
  pivot_longer(cols = -ID1, names_to = "ID2", values_to = "affinity") %>% 
  left_join(gaussian_clusters_nk, by = c("ID1" = "ID")) %>% 
  rename(k1 = k) %>% 
  left_join(gaussian_clusters_nk, by = c("ID2" = "ID")) %>% 
  rename(k2 = k)

patient_lvls <- df_gaussian_affinity_nk %>% 
  select(ID1, k1) %>% 
  distinct() %>% 
  arrange(k1) %>% 
  pull(ID1)

df_gaussian_affinity_nk <- df_gaussian_affinity_nk %>% 
  mutate(ID1 = factor(ID1, levels = patient_lvls),
         ID2 = factor(ID2, levels = patient_lvls),
         affinity = ifelse(ID1 == ID2, NA, affinity))

df_gaussian_affinity_nk %>% 
  mutate(affinity = ifelse(affinity >= 0.015, 0.015, affinity)) %>% 
  ggplot(aes(x = ID1, y = ID2, fill = affinity)) +
  geom_tile() + 
  scale_fill_viridis() + 
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
```



```{r}
mat_gaussian_1 <- rmvnorm(n = 10000, mean = rep(0, 2), sigma = diag(2))
mat_gaussian_2 <- rmvnorm(n = 10000, mean = rep(5, 2), sigma = diag(2))
mat_gaussian <- rbind(mat_gaussian_1, mat_gaussian_2)
colnames(mat_gaussian) <- c("x1", "x2")

ggplot(as_tibble(mat_gaussian), 
       aes(x = x1, y = x2)) + 
  geom_hex() + 
  coord_fixed() + 
  scale_fill_viridis() + 
  theme_bw()
```



