---
title: "Biological pathway modules associated with mouse clusters"
author: "Antoine Beauchamp"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Initialization

```{r packages}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(grid))
suppressPackageStartupMessages(library(gridExtra))
suppressPackageStartupMessages(library(patchwork))
suppressPackageStartupMessages(library(ggalluvial))
suppressPackageStartupMessages(library(pheatmap))
suppressPackageStartupMessages(library(rcartocolor))
suppressPackageStartupMessages(library(RColorBrewer))
suppressPackageStartupMessages(library(viridis))
suppressPackageStartupMessages(library(viridisLite))
```

```{r functions}
source("../../src/analysis.R")
```

```{r params}
#Output directory
output_dir <- "outputs/mouse_cluster_pathways/"
if (!(dir.exists(output_dir))) {dir.create(output_dir, recursive = TRUE)}

# Max number of clusters
nk_max <- 10
```

```{r import-pathway-enrichment}
# Threshold for StringDB 
stringdb_threshold <- 950

# Base directory for pathway data files
pathways_dir <- "/projects/jacob/ClusteringAutism_125Models_Mar2020/Data/Outputs/NeighbourhoodEnrichment_Paper/"
pathways_dir <- file.path(pathways_dir, stringdb_threshold)

# Prefix for pathway data files
pathways_file_prefix <- "NewBader_enrichment_clusterneighbourhood_vs_brain_all"

# Pathway IDs for a prior pathway set
pathway_ids <-c("GAP JUNCTION TRAFFICKING AND REGULATION%REACTOME%R-HSA-157858.1",
                "TRANSMISSION ACROSS CHEMICAL SYNAPSES%REACTOME%R-HSA-112315.5",
                "PROTEIN-PROTEIN INTERACTIONS AT SYNAPSES%REACTOME DATABASE ID RELEASE 71%6794362",
                "ADHERENS JUNCTIONS INTERACTIONS%REACTOME%R-HSA-418990.2",
                "TIGHT JUNCTION INTERACTIONS%REACTOME DATABASE ID RELEASE 71%420029",
                "MAPK FAMILY SIGNALING CASCADES%REACTOME DATABASE ID RELEASE 71%5683057",
                "SIGNALING BY ERBB2%REACTOME DATABASE ID RELEASE 71%1227986",
                "SIGNALING BY ERBB4%REACTOME DATABASE ID RELEASE 71%1236394",
                "SIGNALING BY WNT%REACTOME%R-HSA-195721.5",
                "CA2+ PATHWAY%REACTOME DATABASE ID RELEASE 71%4086398",
                "SIGNALING BY NOTCH%REACTOME DATABASE ID RELEASE 71%157118",
                "MTOR SIGNALLING%REACTOME%R-HSA-165159.5",
                "SIGNALING BY VEGF%REACTOME DATABASE ID RELEASE 71%194138",
                "SIGNALING BY HEDGEHOG%REACTOME DATABASE ID RELEASE 71%5358351",
                "AXON GUIDANCE%REACTOME DATABASE ID RELEASE 71%422475",
                "CHROMATIN ORGANIZATION%REACTOME DATABASE ID RELEASE 71%4839726",
                "GENERIC TRANSCRIPTION PATHWAY%REACTOME%R-HSA-212436.9",
                "GENE EXPRESSION (TRANSCRIPTION)%REACTOME DATABASE ID RELEASE 71%74160",
                "LONG-TERM POTENTIATION%REACTOME DATABASE ID RELEASE 71%9620244",
                "SIGNALING BY GPCR%REACTOME%R-HSA-372790.4")

# List of desired pathway terms
pathway_terms <- c("Chromatin organization",
                   "Gene expression (transcription)",
                   "Generic transcription pathway",
                   "Mapk family signaling cascades",
                   "Mtor signalling",
                   "Protein-protein interactions at synapses",
                   "Signaling by hedgehog",
                   "Signaling by notch",
                   "Signaling by wnt",
                   "Transmission across chemical synapses")

# Import mouse pathway enrichment files
# Iterate over cluster solutions
list_pathways <- vector(mode = "list", length = nk_max-1)
names(list_pathways) <- 2:nk_max
for (nk in 2:nk_max) {
  
  # Iterate over cluster number
  list_pathways[[as.character(nk)]] <- vector(mode = "list", length = nk)
  for (k in 1:nk) {
    pathways_file <- paste(pathways_file_prefix, nk, k, stringdb_threshold, sep = "_")
    pathways_file <- paste0(pathways_file, ".csv")
    pathways_file <- file.path(pathways_dir, pathways_file)
    list_pathways[[as.character(nk)]][[k]] <- read_csv(pathways_file, show_col_types = FALSE)  
    
    # Fix mouse enrichment p-values and q-values
    # REMOVE IF FIXED UPSTREAM
    list_pathways[[as.character(nk)]][[k]] <- list_pathways[[as.character(nk)]][[k]] %>% 
      mutate(P.Value = P.Value/2,
             adj.P.Val = p.adjust(P.Value, method = "fdr"),
             NLQ = -log10(adj.P.Val),
             qval_log = -log10(adj.P.Val))
  }
  
  # Combine clusters per solution
  list_pathways[[as.character(nk)]] <- list_pathways[[as.character(nk)]] %>% 
    reduce(.f = bind_rows) %>% 
    rename(pathway = Title,
           k = cluster) %>% 
    mutate(nk = nk) %>% 
    unite(col = "cluster_id", nk, k, 
          sep = "-", remove = FALSE)
  
}
```


# Pathway enrichment polar plots

```{r generate-polar-plots}
# Polar plot labels
polar_plot_lvls <- c(" ", pathway_terms)

# Create polar plot data frames
list_polar_plots <- vector(mode = "list", length = length(list_pathways))
names(list_polar_plots) <- names(list_pathways)
for (i in 1:length(list_polar_plots)) {
  
  df_pathways_nk <- list_pathways[[i]] %>% 
    filter(pathway %in% pathway_terms)
  
  df_polar_plot_tmp <- df_pathways_nk %>% 
    mutate(pathway = as.character(pathway)) %>% 
    filter(pathway == polar_plot_lvls[length(polar_plot_lvls)]) %>% 
    mutate(pathway = " ")
  
  list_polar_plots[[i]] <- df_pathways_nk %>% 
    bind_rows(df_polar_plot_tmp) %>% 
    mutate(pathway = factor(pathway, levels = polar_plot_lvls))
  
}

# Polar plot graphical parameters
nspokes <- length(polar_plot_lvls) - 1
theta_start <- -(2*pi)/nspokes
breaks_start <- 0
E_breaks_step <- 5
q_breaks_step <- 1

# Polar plot palette
polar_palette <- carto_pal(n = 12, name = "Bold")

# Enrichment and q-value clamping thresholds
E_threshold <- 30
q_threshold <- 5

# Iterate over clusters
for (i in 1:length(list_polar_plots)) {
  
  # Radial labels
  E_max <- ceiling(max(list_polar_plots[[i]][["E"]]))
  q_max <- ceiling(max(list_polar_plots[[i]][["qval_log"]]))
  
  E_max <- ifelse(E_max > E_threshold, E_threshold, E_max)
  q_max <- ifelse(q_max > q_threshold, q_threshold, q_max)
  
  E_breaks_end <- E_max - E_max %% E_breaks_step + E_breaks_step
  q_breaks_end <- q_max - q_max %% q_breaks_step + q_breaks_step
  
  E_polar_breaks <- seq(breaks_start, E_breaks_end, by = E_breaks_step)
  q_polar_breaks <- seq(breaks_start, q_breaks_end, by = q_breaks_step)
  
  E_polar_labels <- tibble(pathway = polar_plot_lvls[2],
                           breaks = E_polar_breaks)
  
  q_polar_labels <- tibble(pathway = polar_plot_lvls[2],
                           breaks = q_polar_breaks)
  
  df_polar_plot <- list_polar_plots[[i]] %>% 
    select(k, pathway, E, qval_log) %>% 
    mutate(E = ifelse(E > E_threshold, E_threshold, E),
           qval_log = ifelse(qval_log > q_threshold, q_threshold, qval_log),
           k = factor(k))
  
  E_polar_plot <- ggplot(df_polar_plot, 
                         aes(x = pathway, 
                             ymin = 0, ymax = E, 
                             group = k, 
                             fill = k, 
                             col = k)) + 
    geom_ribbon(alpha = 0.2) + 
    geom_text(data = E_polar_labels,
              inherit.aes = FALSE,
              mapping = aes(x = pathway,
                            y = breaks,
                            label = breaks)) +
    coord_radar(start = theta_start, clip = "off") + 
    scale_x_discrete(expand = expansion()) + 
    scale_y_continuous(breaks = E_polar_breaks) +
    scale_fill_manual(values = polar_palette[1:(i+1)]) +
    scale_colour_manual(values = polar_palette[1:(i+1)]) +
    labs(x = NULL,
         fill = "k",
         col = "k") + 
    theme_bw()
  
  q_polar_plot <- ggplot(df_polar_plot, 
                         aes(x = pathway, 
                             ymin = 0, ymax = qval_log, 
                             group = k, 
                             fill = k, 
                             col = k)) + 
    geom_ribbon(alpha = 0.2) + 
    geom_text(data = q_polar_labels,
              inherit.aes = FALSE,
              mapping = aes(x = pathway,
                            y = breaks,
                            label = breaks)) +
    coord_radar(start = theta_start, clip = "off") + 
    scale_x_discrete(expand = expansion()) + 
    scale_y_continuous(breaks = q_polar_breaks) +
    scale_fill_manual(values = polar_palette[1:(i+1)]) +
    scale_colour_manual(values = polar_palette[1:(i+1)]) +
    labs(x = NULL,
         fill = "k",
         col = "k") + 
    theme_bw()
  
  polar_plot <- (E_polar_plot | q_polar_plot) +
    plot_layout(guides = "collect") &
    theme(axis.ticks.x = element_blank(),
          axis.line.x = element_blank(),
          axis.text.x = element_text(size = 6),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.title.y = element_blank(),
          panel.border = element_blank(),
          legend.position = "bottom")
  
  # Write to file
  outfile <- paste0("pathways_polar_nk_", names(list_polar_plots)[i], ".pdf")
  outfile <- file.path(output_dir, outfile)
  pdf(file = outfile,
      width = unit(12, "in"),
      height = unit(6, "in"))
  print(polar_plot)
  dev.off()
  
}
```

# Pathway enrichment Sankey plots

```{r import-clusters}
# Import mouse clusters
# cluster_file <- file.path(cluster_dir, "clusters.csv")
cluster_file <- "../../data/mouse/derivatives/v2/107/clusters/clusters.csv"
df_clusters <- read_csv(cluster_file, show_col_types = FALSE)
colnames(df_clusters) <- c("ID", str_c("nk", 2:nk_max))

# Convert mouse clusters to long format
df_clusters_long <- df_clusters %>% 
  pivot_longer(cols = -ID, names_to = "nk_name", values_to = "k") %>% 
  mutate(nk = str_remove(nk_name, "nk"),
         nk = as.numeric(nk),
         nk = factor(nk),
         k = factor(k))
```

```{r generate-sankey-plots}
# Iterate over pathways
for (i in 1:length(pathway_ids)) {
  
  # Filter for given pathway in each cluster
  list_pathway_i <- vector(mode = "list", length = length(list_pathways))
  for (j in 1:length(list_pathway_i)) {
    list_pathway_i[[j]] <- list_pathways[[j]] %>% 
      filter(ID == pathway_ids[i])
  }
  
  # Reduce all cluster enrichment for pathway i into one df
  df_pathway_i <- list_pathway_i %>% 
    bind_rows() %>% 
    select(-ID) %>% 
    mutate(nk = factor(nk, levels = 2:nk_max),
           k = factor(k, levels = 1:nk_max),
           E = ifelse(E > E_threshold, E_threshold, E)) %>% 
    right_join(df_clusters_long, by = c("nk", "k"))
  
  # Generate Sankey plot
  p_sankey_pathway <- ggplot(df_pathway_i, 
                             aes(x = nk, 
                                 stratum = k,
                                 fill = E,
                                 alluvium = ID)) + 
    geom_flow(stat = "alluvium", aes.flow = "forward") + 
    geom_stratum() + 
    scale_fill_gradientn(colors = brewer.pal(n = 9, name = "OrRd")[3:9],
                         limits = c(0, E_threshold)) + 
    # limits = c(0.3, 3),
    # breaks = seq(0, 3, by = 0.5)) +
    labs(x = "Number of clusters",
         y = "Number of models",
         fill = "Enrichment (clamped)",
         title = unique(df_pathway_i[["pathway"]])) + 
    theme_bw() + 
    theme(panel.grid.major.x = element_blank())
  
  # Output file name
  outfile <- "pathways_sankey_" %>%
    str_c(df_pathway_i %>% 
            pull(pathway) %>% 
            unique() %>% 
            str_to_lower() %>% 
            str_remove_all("[()]") %>% 
            str_replace_all(" ", "_")) %>% 
    str_c(".pdf")
  outfile <- file.path(output_dir, outfile)
  
  # Export plot
  pdf(file = outfile,
      width = unit(10, "inch"),
      height = unit(5, "inch"))
  print(p_sankey_pathway)
  dev.off()
  
}
```

# Pathway enrichment heatmaps

## Cluster significance by permutation testing

```{r import-perms}
# Permutation directory and files
# permutation_dir <- file.path(pipeline_dir, params_id, "permutations", "similarity")
permutation_dir <- "../../data/cross_species/v2/405/permutations/similarity/"
permutation_files <- list.files(permutation_dir, full.names = TRUE)

# Permutation IDs
permutation_ids <- permutation_files %>% 
  basename() %>% 
  str_extract("[0-9]+") %>% 
  as.numeric() %>% 
  unique() %>% 
  sort()

# Jacobians to use
jacobians <- c("absolute", "relative")
#  jacobians <- "relative"

# Number of permutations
np <- length(permutation_ids)

# Iterate over jacobians
list_permutations <- vector(mode = "list", length = length(jacobians))
names(list_permutations) <- jacobians
for (j in jacobians) {
  
  # Iterate over permutations
  df_sim <- tibble()
  for (p in 1:np) {
    
    # Permutation similarity data
    infile <- permutation_files %>% 
      str_subset(str_c("permutation_", permutation_ids[p], "_", j))
    
    # Import permutation similarity data
    df_sim_tmp <- infile %>% 
      read_csv(show_col_types = FALSE) %>% 
      mutate(human_nk = human_img %>% basename() %>% str_extract("_nk_[0-9]+") %>% str_extract("[0-9]+") %>% as.numeric(),
             human_k = human_img %>% basename() %>% str_extract("_k_[0-9]+") %>% str_extract("[0-9]+") %>% as.numeric(),
             mouse_nk = mouse_img %>% basename() %>% str_extract("_nk_[0-9]+") %>% str_extract("[0-9]+") %>% as.numeric(),
             mouse_k = mouse_img %>% basename() %>% str_extract("_k_[0-9]+") %>% str_extract("[0-9]+") %>% as.numeric()) %>% 
      unite(col = "human_cluster_id", human_nk, human_k, sep = "-", remove = FALSE) %>% 
      unite(col = "mouse_cluster_id", mouse_nk, mouse_k, sep = "-", remove = FALSE) %>% 
      mutate(permutation = permutation_ids[p]) 
    
    # Collate permutation data
    df_sim <- bind_rows(df_sim, df_sim_tmp)
  }
  
  # Include jacobians information
  df_sim <- df_sim %>% 
    mutate(jacobians = j)
  
  list_permutations[[j]] <- df_sim
  
}

# Reduce permutations data frames into one
df_permutations <- reduce(.x = list_permutations, .f = bind_rows)

# Average similarity values for both jacobians
df_permutations <- df_permutations %>% 
  group_by(permutation, 
           human_cluster_id, human_nk, human_k, 
           mouse_cluster_id, mouse_nk, mouse_k) %>% 
  summarise(similarity = mean(similarity), .groups = "drop")
```

```{r import-sim}
# Cluster similarity files
# similarity_dir <- file.path(pipeline_dir, params_id, "similarity")
similarity_dir <- "../../data/cross_species/v2/405/similarity/"
similarity_files <- list.files(similarity_dir, full.names = TRUE)

# Import absolute and relative similarity files
similarity <- vector(mode = "list", length = length(jacobians))
names(similarity) <- jacobians
for (j in jacobians) {
  similarity[[j]] <- similarity_files %>% 
    str_subset(j) %>% 
    read_csv(show_col_types = FALSE) %>% 
    mutate(human_nk = human_img %>% basename() %>% str_extract("_nk_[0-9]+") %>% str_extract("[0-9]+") %>% as.numeric(),
           human_k = human_img %>% basename() %>% str_extract("_k_[0-9]+") %>% str_extract("[0-9]+") %>% as.numeric(),
           mouse_nk = mouse_img %>% basename() %>% str_extract("_nk_[0-9]+") %>% str_extract("[0-9]+") %>% as.numeric(),
           mouse_k = mouse_img %>% basename() %>% str_extract("_k_[0-9]+") %>% str_extract("[0-9]+") %>% as.numeric()) %>% 
    unite(col = "human_cluster_id", human_nk, human_k, sep = "-", remove = FALSE) %>% 
    unite(col = "mouse_cluster_id", mouse_nk, mouse_k, sep = "-", remove = FALSE) %>% 
    mutate(jacobians = j)
}

# Reduce similarity data frames into one
df_similarity <- reduce(.x = similarity, .f = bind_rows)

# Average similarity values for both jacobians
df_similarity <- df_similarity %>% 
  group_by(human_cluster_id, human_nk, human_k, 
           mouse_cluster_id, mouse_nk, mouse_k) %>% 
  summarise(similarity = mean(similarity), .groups = "drop")
```

```{r compute-pvals}
# Mouse and human max nk
human_nk_max <- max(df_similarity[["human_nk"]])
mouse_nk_max <- max(df_similarity[["mouse_nk"]])

# Iterate along nk diagonal +/- 1
df_sim_pvals <- tibble()
for (h_nk in 2:human_nk_max) {
  for (m_nk in (h_nk-1):(h_nk+1)){
    
    if ((m_nk > 1) & (m_nk <= mouse_nk_max)) {
      
      df_sim_nk <- df_similarity %>% 
        select(human_cluster_id, human_nk, human_k,
               mouse_cluster_id, mouse_nk, mouse_k,
               similarity) %>% 
        filter(human_nk == h_nk,
               mouse_nk == m_nk) %>% 
        mutate(pval = 0)
      
      sim_perm_nk <- df_permutations %>% 
        filter(human_nk == h_nk,
               mouse_nk == m_nk) %>% 
        pull(similarity) %>% 
        sort()
      
      for (i in 1:nrow(df_sim_nk)) {
        ntail <- sum(sim_perm_nk >= df_sim_nk[[i, "similarity"]])
        df_sim_nk[[i, "pval"]] <- ntail/length(sim_perm_nk)
      }
      
      df_sim_pvals <- bind_rows(df_sim_pvals, df_sim_nk)
      
    }
  }
}

# Evaluate significance
alpha <- 0.05
df_sim_pvals <- df_sim_pvals %>% 
  mutate(significant = ifelse(pval < 0.05, 1, 0))
```


## Generate heatmap

```{r pathways-heatmap-proc}
# Reduce all pathway data frames into one
df_pathways_all <- bind_rows(list_pathways)

# Filter for cluster solutions with nk >= 4
df_pathways_nk_gt4 <- df_pathways_all %>% 
  filter(nk >= 4)

# Cluster order 
cluster_lvls <- df_pathways_nk_gt4 %>% 
  select(nk, k, cluster_id) %>% 
  distinct() %>% 
  arrange(nk, k) %>% 
  pull(cluster_id)

# Filter for desired pathways
# Compute normalized enrichment per pathway
df_pathways_nk_gt4_subset <- df_pathways_nk_gt4 %>% 
  filter(pathway %in% pathway_terms) %>% 
  mutate(cluster_id = factor(cluster_id, levels = cluster_lvls),
         pathway = factor(pathway, levels = pathway_terms)) %>% 
  group_by(pathway) %>% 
  mutate(E_norm = E/max(E)) %>% 
  ungroup()

# Convert pathway data frame into matrix 
mat_pathways_nk_gt4_subset <- df_pathways_nk_gt4_subset %>% 
  select(pathway, cluster_id, E_norm) %>% 
  pivot_wider(id_cols = pathway, 
              names_from = cluster_id, 
              values_from = E_norm) %>% 
  column_to_rownames("pathway") %>% 
  as.matrix()
```

```{r pathways-heatmap-clust}
# Generate clustered heatmap
pheatmap_out <- pheatmap(mat = mat_pathways_nk_gt4_subset,
                         cluster_col = FALSE,
                         cluster_rows = TRUE,
                         clustering_distance_rows = "euclidean",
                         cutree_rows = 4, 
                         silent = TRUE)

# Extract pathway clustering outputs
pheatmap_pathway_lvls <- pheatmap_out[["tree_row"]][["labels"]]
pheatmap_pathway_order <- pheatmap_out[["tree_row"]][["order"]]
pathway_lvls_clustered <- pheatmap_pathway_lvls[pheatmap_pathway_order]

# Clustering order for pathway labels
# pathway_label_lvls_clustered <- df_pathway_labels %>% 
#   mutate(pathway = factor(pathway, levels = pathway_lvls_clustered)) %>% 
#   arrange(pathway) %>% 
#   pull(label)

# Level at which to cut the dendrogram
hclust_kcut <- 4

# Obtain pathway cluster order at selected solution
df_pathway_hclust <- cutree(pheatmap_out[["tree_row"]], k = hclust_kcut) %>% 
  enframe(name = "pathway", value = "pathway_cluster") 

# Relevel pathways to follow dendrogram order
df_pathway_cluster_lvls <- df_pathway_hclust %>% 
  mutate(pathway = factor(pathway, levels = pathway_lvls_clustered)) %>% 
  arrange(pathway) %>% 
  select(pathway_cluster) %>% 
  distinct() %>% 
  mutate(pathway_cluster_new = 1:nrow(.))

df_pathway_hclust <- df_pathway_hclust %>% 
  left_join(df_pathway_cluster_lvls, by = "pathway_cluster") %>% 
  select(-pathway_cluster, pathway_cluster = pathway_cluster_new)

# Extract dendrogram from pheatmap object
pathways_dendrogram_grob <- pheatmap_out[["gtable"]][["grobs"]][[1]]
pathways_dendrogram_grob[["gp"]] <- gpar(lwd = 0.75)
```

```{r pathways-heatmap-human-match}
df_sig_nk <- df_sim_pvals %>%
  group_by(mouse_nk, mouse_k) %>% 
  summarise(nsignificant = sum(significant),
            .groups = "drop") %>%
  mutate(has_match = ifelse(nsignificant > 0, TRUE, FALSE),
         dummy = "y") %>% 
  select(nk = mouse_nk, k = mouse_k, has_match, dummy) %>% 
  filter(nk >= 4) %>% 
  unite(col = "cluster_id", "nk", "k", 
        sep = "-", remove = FALSE) %>% 
  mutate(nk = factor(nk, levels = 4:nk_max),
         k = factor(k, levels = 1:nk_max),
         cluster_id = factor(cluster_id, levels = cluster_lvls))

human_match_heatmap <- ggplot(df_sig_nk,
                              mapping = aes(x = k, 
                                            y = dummy, 
                                            fill = has_match)) + 
  geom_tile(col = "grey50") +
  facet_grid(.~nk, scales = "free", space = "free") + 
  scale_x_discrete(expand = expansion()) + 
  scale_y_discrete(expand = expansion()) +
  scale_fill_manual(values = brewer.pal(n = 9, name = "PuBu")[c(1, 9)], 
                    labels = c("No", "Yes")) + 
  labs(fill = "Human match") + 
  theme_bw() +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank())
```

```{r pathways-heatmap-enrichment}
# Join pathway enrichment information with pathway clusters
df_pathways_heatmap <- df_pathways_nk_gt4_subset %>% 
  left_join(df_pathway_hclust, by = "pathway") %>% 
  mutate(pathway = factor(pathway, levels = pathway_lvls_clustered))

# Clamped enrichment
E_threshold <- 45
df_pathways_heatmap <- df_pathways_heatmap %>% 
  mutate(E_clamped = ifelse(E > E_threshold, E_threshold, E))

# Heatmap palette
heatmap_palette_cols <- c("white", brewer.pal(n = 9, name = "OrRd")[3:9])
heatmap_palette <- colorRampPalette(colors = heatmap_palette_cols)(255)

# Generate heatmap
pathways_heatmap <- ggplot(df_pathways_heatmap, 
                           aes(x = factor(k), 
                               y = fct_rev(pathway), 
                               fill = E_clamped)) + 
  geom_tile(col = "grey60") +
  facet_grid(pathway_cluster~nk, scales = "free", space = "free") + 
  scale_x_discrete(expand = expansion()) + 
  scale_y_discrete(expand = expansion(), 
                   position = "right") + 
  scale_fill_gradientn(colors = heatmap_palette) +
  labs(x = "Mouse cluster",
       y = "Biological pathway module",
       fill = "Enrichment") +
  theme_bw() +
  theme(strip.background = element_blank(),
        strip.text = element_blank())
```

```{r pathways-heatmap-patchwork}
heatmap_patchwork <- (human_match_heatmap / pathways_heatmap) +
  plot_layout(heights = c(0.06, 0.94), guides = "collect") &
  theme(plot.margin = margin(l = 0))

heatmap_patchwork_grob <- patchworkGrob(heatmap_patchwork)
```

```{r generate-pathways-heatmap}
# Plot layout
pathways_layout <- rbind(c(01, 02),
                         c(03, 02),
                         c(04, 02))

# Plot grid dimensions
pathways_widths <- c(0.05, 0.95)
pathways_heights <- c(0.11, 0.80, 0.09)

empty_rect_grob <- rectGrob(gp = gpar(fill = NA))

# Plot arrange grob
pathways_grob <- arrangeGrob(zeroGrob(),
                             heatmap_patchwork_grob,
                             pathways_dendrogram_grob,
                             zeroGrob(),
                             layout_matrix = pathways_layout,
                             widths = unit(pathways_widths, "npc"),
                             heights = unit(pathways_heights, "npc"))

# Plot dimensions
pathways_width_in <- 12
pathways_height_in <- 4

# Plot dimensions with units
pathways_width_in <- unit(pathways_width_in, "in")
pathways_height_in <- unit(pathways_height_in, "in")

# Export plot
outfile <- "pathways_heatmap.pdf"
outfile <- file.path(output_dir, outfile)
pdf(file = outfile,
    width = pathways_width_in,
    height = pathways_height_in)
grid.draw(pathways_grob)
dev.off()
```

```{r generate-pathways-heatmap-norm}
# Generate heatmap
pathways_heatmap_norm <- ggplot(df_pathways_heatmap, 
                                aes(x = factor(k), 
                                    y = fct_rev(pathway), 
                                    fill = E_norm)) + 
  geom_tile(col = "grey60") +
  facet_grid(pathway_cluster~nk, scales = "free", space = "free") + 
  scale_x_discrete(expand = expansion()) + 
  scale_y_discrete(expand = expansion(), 
                   position = "right") + 
  scale_fill_gradientn(colors = heatmap_palette) +
  labs(x = "Mouse cluster",
       y = "Biological pathway module",
       fill = "Enrichment") +
  theme_bw() +
  theme(strip.background.y = element_blank(),
        strip.text.y = element_blank(),
        plot.margin = margin(l = 0))

heatmap_patchwork_norm <- (human_match_heatmap / pathways_heatmap_norm) +
  plot_layout(heights = c(0.06, 0.94), guides = "collect") &
  theme(plot.margin = margin(l = 0))

heatmap_patchwork_norm_grob <- patchworkGrob(heatmap_patchwork_norm)

# Plot arrange grob
pathways_norm_grob <- arrangeGrob(zeroGrob(),
                                  heatmap_patchwork_norm_grob,
                                  pathways_dendrogram_grob,
                                  zeroGrob(),
                                  layout_matrix = pathways_layout,
                                  widths = unit(pathways_widths, "npc"),
                                  heights = unit(pathways_heights, "npc"))

outfile <- "pathways_heatmap_norm.pdf"
outfile <- file.path(output_dir, outfile)

pdf(file = outfile,
    width = pathways_width_in,
    height = pathways_height_in)
grid.draw(pathways_norm_grob)
dev.off()
```

# All pathways

What is the proper way to look at this? What am I looking for? For each cluster, what are the most enriched pathways? Maybe start with one cluster to figure out what it is that I want to do. 

```{r}
df_pathways_all_nk_geq4 <- df_pathways_all %>% 
  filter(nk >= 4)

nhighest <- 20

list_pathways_all_topn <- vector(mode = "list", length = length(cluster_lvls))
for (i in 1:length(list_pathways_all_topn)) {
  list_pathways_all_topn[[i]] <- df_pathways_all_nk_geq4 %>% 
    filter(cluster_id == cluster_lvls[i]) %>% 
    arrange(rank) %>% 
    head(n = nhighest)
}

df_pathways_all_topn <- bind_rows(list_pathways_all_topn)

top_pathway_lvls <- df_pathways_all_topn %>% 
  group_by(pathway) %>% 
  count() %>% 
  arrange(desc(n)) %>% 
  pull(pathway)

df_pathways_all_topn <- df_pathways_all_topn %>%
  mutate(pathway = factor(pathway, levels = top_pathway_lvls),
         k = factor(k),
         nk = factor(nk),
         cluster_id = factor(cluster_id, levels = cluster_lvls),
         NLQ = ifelse(NLQ > 20, 20, NLQ))

heatmap_topn_palette_cols <- carto_pal(n = 7, "BurgYl")
heatmap_topn_palette <- colorRampPalette(colors = heatmap_topn_palette_cols)(255)

pathways_heatmap_topn <- ggplot(df_pathways_all_topn,
                                aes(x = k, 
                                    y = fct_rev(pathway), 
                                    fill = NLQ)) + 
  geom_tile(col = "grey60") + 
  facet_grid(.~nk, scales = "free", space = "free") +
  scale_x_discrete(expand = expansion()) +
  scale_y_discrete(expand = expansion()) +
  scale_fill_gradientn(colors = heatmap_topn_palette,
                       limits = c(1.3, 20)) +
  labs(y = "Biological pathway module") + 
  theme_bw() +
  theme(axis.text = element_text(size = 6))

outfile <- paste("pathways_heatmap_top", nhighest, sep = "_")
outfile <- paste0(outfile, ".pdf")
outfile <- file.path(output_dir, outfile)
pdf(file = outfile,
    width = unit(10, "in"),
    height = unit(20, "in"))
print(pathways_heatmap_topn)
dev.off()
```

This is a lot of pathways. Looking at the top 20 pathways in each cluster returns 165 pathways. I need to filter this down in some way. Two things I can think of: 
1. Look only at those clusters that have human matches. 
2. Some sort of thresholding based on the number of genes in a given pathway. Presumably the more general pathways will be associated with more genes. 


This should be contained in one of the hypergeometric distribution parameters. 
N is the background set, i.e. the set of all genes we're testing against. 
n is the number of genes in the cluster neighbourhood that are also in the background set. 
B is the number of genes in the target set, i.e. the set of genes associated with the given molecular pathway. I think this is the value I'm looking for. 
b is the number of genes in the cluster neighbourhood that are also in the target set. 

```{r}
df_pathways_all %>% 
  select(pathway, B) %>% 
  distinct()
```

```{r}
B <- sort(df_pathways_all_nk_geq4[["B"]])
B_max <- max(B)
B_min <- 0
B_n <- 1e3
B_seq <- seq(B_min, B_max, length.out = B_n)
B_ecdf <- numeric(length(B_seq))
for (i in 1:B_n) {B_ecdf[i] <- sum(B <= B_seq[i])/length(B)}

tibble(B = B_seq,
       cdf = B_ecdf) %>% 
  ggplot(aes(x = B, y = cdf)) + 
  geom_line() + 
  geom_point(size = 0.5)
```
```{r}
B_threshold_p <- 0.9
B_threshold <- ceiling(B_seq[min(which(B_ecdf >= B_threshold_p))])

nhighest <- 20

list_pathways_all_topn <- vector(mode = "list", length = length(cluster_lvls))
for (i in 1:length(list_pathways_all_topn)) {
  list_pathways_all_topn[[i]] <- df_pathways_all_nk_geq4 %>%
    filter(cluster_id == cluster_lvls[i],
           B <= B_threshold) %>% 
    arrange(rank) %>% 
    head(n = nhighest)
}

df_pathways_all_topn <- bind_rows(list_pathways_all_topn)

top_pathway_lvls <- df_pathways_all_topn %>% 
  group_by(pathway) %>% 
  count() %>% 
  arrange(desc(n)) %>% 
  pull(pathway)

df_pathways_all_topn <- df_pathways_all_topn %>%
  mutate(pathway = factor(pathway, levels = top_pathway_lvls),
         k = factor(k),
         nk = factor(nk),
         cluster_id = factor(cluster_id, levels = cluster_lvls),
         NLQ = ifelse(NLQ > 20, 20, NLQ))

heatmap_topn_palette_cols <- carto_pal(n = 7, "BurgYl")
heatmap_topn_palette <- colorRampPalette(colors = heatmap_topn_palette_cols)(255)

pathways_heatmap_topn <- ggplot(df_pathways_all_topn,
                                aes(x = k, 
                                    y = fct_rev(pathway), 
                                    fill = NLQ)) + 
  geom_tile(col = "grey60") + 
  facet_grid(.~nk, scales = "free", space = "free") +
  scale_x_discrete(expand = expansion()) +
  scale_y_discrete(expand = expansion()) +
  scale_fill_gradientn(colors = heatmap_topn_palette,
                       limits = c(1.3, 20)) +
  labs(y = "Biological pathway module") + 
  theme_bw() +
  theme(axis.text = element_text(size = 6))

outfile <- paste("pathways_heatmap_top", nhighest, "B_threshold", B_threshold_p, sep = "_")
outfile <- paste0(outfile, ".pdf")
outfile <- file.path(output_dir, outfile)
pdf(file = outfile,
    width = unit(10, "in"),
    height = unit(20, "in"))
print(pathways_heatmap_topn)
dev.off()
```

